---
title: "Faraway -Linear Model - 2 ed - Chapter 1 "
author: "Andrea Pierucci"
date: "01 May 2020"
output: html_document
---
1)

```{r}
library(faraway)
attach(teengamb)                
data(teengamb)
head(teengamb)
summary(teengamb)
```

Sex is a categorical variable

```{r}
teengamb$sex <-as.factor(teengamb$sex) 
```

A numerical summary divided by sex is useful for understanding patterns if present
```{r}

by(teengamb, teengamb$sex, summary)
pairs(teengamb) 
cor(teengamb[,-1])
```
We remove the qualitative variable were correlations cannot be evaluated. 

Female 
```{r}
cor(teengamb[teengamb$sex==1,-1])  
```
Male
```{R}
cor(teengamb[teengamb$sex==0,-1]) 
```

Graphical summary 
```{r}
plot(gamble~sex, teengamb) 
```
We can identify outliers
```{r}
plot(income~sex)     
plot(gamble~income, pch=unclass(sex))  

```


2)

```{r}
data(uswages)
attach(uswages)                
head(uswages)
summary(uswages)
```

Change categorical variables

```{r}
uswages$race <-  as.factor(uswages$race) 
uswages$smsa <-  as.factor(uswages$smsa) 
uswages$ne <-  as.factor(uswages$ne) 
uswages$mw <-  as.factor(uswages$mw) 
uswages$so <-  as.factor(uswages$so) 
uswages$we <-  as.factor(uswages$we) 
uswages$pt <-  as.factor(uswages$pt) 
```

A numerical summary divided by race is useful for understanding patterns if present
```{r}

by(uswages, uswages$race, summary)
pairs(uswages[,-c(4:10)]) 
cor(uswages[,-c(4:10)])
```
We remove the qualitative variables were correlations cannot be evaluated. 

Black 
```{r}
cor(uswages[uswages$race==1,-c(4:10)])  
```
White
```{R}
cor(uswages[uswages$race==0,-c(4:10)]) 
```

Graphical summary 
```{r}
plot(educ~race, uswages)
plot(exper~race, uswages)
plot(wage~race, uswages)

```
We can identify outliers
```{r}
    
plot(wage~educ, pch=unclass(race)) 
plot(wage~exper, pch=unclass(race)) 
plot(educ~exper, pch=unclass(race)) 

```

3)
```{r}
data(prostate)
attach(prostate)                
head(prostate)
summary(prostate)
```

Change categorical variables
```{r}
prostate$svi <-  as.factor(prostate$svi) 
```
A numerical summary divided by svi is useful for understanding patterns if present
```{r}
by(prostate, prostate$svi, summary)
pairs(prostate[,-5]) 
cor(prostate[,-5])
```
We remove the qualitative variables were correlations cannot be evaluated. 

With seminal vesicle invasion 
```{r}
cor(prostate[prostate$svi==1,-5])  
```
Without seminal vesicle invasion 
```{r}
cor(prostate[prostate$svi==0,-5]) 
```
Graphical summary 
```{r}
plot(age~svi, prostate)
plot(lweight~svi, prostate)
plot(lweight~age, prostate)
plot(lcavol~age, prostate)
```

4)
```{r}
data(sat)
attach(sat)
summary(sat)
pairs(sat)
cor(sat)
```
High correlation between salary vs expenditure, verbal vs math, verbal vs total, math vs total

we can observe those strong correlation garaphically 
```{r}
plot(expend~salary)      
plot(verbal~math)      
plot(verbal~total)      
plot(math~total)      
```


5)
```{r}
data(divusa)
attach(divusa)
head(divusa)
```
Numerical summary 
```{r}
summary(divusa)
cor(divusa)
```

Graphical summary 
```{r}
pairs(divusa) 
divusa$year<-as.character(divusa$year)
plot(marriage~year,divusa)     
plot(femlab~year,divusa) 
```
